---
title: "Homework 2"
author: "Zeyan Huang"
date: "`r format(Sys.time(), '%B %d, %Y')`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Instructions

**Exercises:**  1-3 (Pgs. 49-50); 2-4 (Pg. 51); 2,4 (Pg. 54); 1-4 (Pg. 58)

**Submission:** Submit via an electronic document on Sakai. Must be submitted as a HTML file generated in RStudio. All assigned problems are chosen according to the textbook *R for Data Science*. You do not need R code to answer every question. If you answer without using R code, delete the code chunk. If the question requires R code, make sure you display R code. If the question requires a figure, make sure you display a figure. A lot of the questions can be answered in written response, but require R code and/or figures for understanding and explaining. 

```{r, include=FALSE}
library(tidyverse)
library(ggplot2)
library(nycflights13)
```

# Chapter 3 (Pgs. 49-50)

##  Exercise 1
```{r}
#
filter(flights, arr_delay >= 120)
filter(flights, dest == 'HOU' | dest == 'IAH')
filter(flights, carrier == 'UA'| carrier == 'AA' | carrier == 'DL')
filter(flights, month >= 7, month <= 9)
filter(flights, dep_delay <= 0, arr_delay > 120)
filter(flights, dep_delay >= 60, dep_delay - arr_delay > 30)
filter(flights, dep_time <= 600 | dep_time == 2400)
```

##  Exercise 2
```{r}
#
filter(flights, between(dep_time, 0, 600))
```

between() function is used to check that whether a numeric value falls in a specific range or not. It can help us simplify the code we need when we try to find out some data in a specific range like what we did in exercise 1 d, e, f, g.

##  Exercise 3
```{r}
#
filter(flights, is.na(dep_time))
```

There are 8,225 flights have a missing dep_time. They also have missing dep_delay, arr_time, arr_delay, and air_time. These rows represent that we don't know the exact time of departure and arrival of these flights, they maybe are flights had been cancelled.

# Chapter 3 (Pg. 51)

##  Exercise 2
```{r}
#
arrange(flights, desc(dep_delay))
arrange(flights,dep_delay)
```

The most delayed flights is HA 51, which was scheduled to leave on January 09, 2013 09:00. And the flight left earliest is B6 97, which was scheduled to leave on December 07, 2013 21:23.

##  Exercise 3
```{r}
#
mutate_flights_speed<-mutate(flights,
                     speed = distance/air_time)
select(arrange(mutate_flights_speed, desc(speed)), speed, everything())
```

##  Exercise 4
```{r}
#
arrange(flights, distance)
arrange(flights, desc(distance))
```

The flight US 1632 traveled the shortest and the flight HA 51 traveled the longest.

# Chapter 3 (Pg. 54)

##  Exercise 2
```{r}
#
select(flights, dep_time, dep_time, everything())
```

The variable which has been mention in a select() call multiple times will only show once in the output.

##  Exercise 4
```{r}
#
select(flights, contains("TIME"))
select(flights, contains("TIME", ignore.case = FALSE))
```

The select helper ignore the difference between different case by default. If we want to change the default, we need to add 'ignore.case = False' by manual since it set to be TRUE by default.

# Chapter 3 (Pg. 58)

##  Exercise 1
```{r}
#
mutate_flights_continuous_time<-mutate(flights,
                     dep_hour = dep_time%/%100,
                     dep_minute = dep_time%%100,
                     dep_time_continuous = (dep_hour*60 + dep_minute)%%1440,
                     sched_dep_hour = sched_dep_time%/%100,
                     sched_dep_minute = sched_dep_time%%100,
                     sched_dep_time_continuous = (sched_dep_hour*60 + sched_dep_minute)%%1440)
select(mutate_flights_continuous_time, dep_time_continuous, sched_dep_time_continuous, everything())
                     
```

##  Exercise 2
```{r}
#
transmute(flights,
            air_time,
            arr_time - dep_time)
```


```{r}
flights_airtime <-
  mutate(flights,
    dep_time = (dep_time %/% 100 * 60 + dep_time %% 100) %% 1440,
    arr_time = (arr_time %/% 100 * 60 + arr_time %% 100) %% 1440,
    air_time_diff = air_time - arr_time + dep_time
  )
select(flights_airtime, air_time_diff, everything())
```
Actually, I expect to see this two number should be equal for each flight, but they do not. I think what we need to do is to convert both the arr_time and dep_time into continuous numbers, like what I did in exercise 1. Then I tried it in continuous number but it still turns out with some data with air_time != arr_time - dep_time, which means this relation may not be true.

##  Exercise 3
```{r}
#
select(arrange(flights, desc(dep_delay)), dep_time, sched_dep_time, dep_delay)

flights_deptime <-
  mutate(flights,
    dep_time_continuous = (dep_time %/% 100 * 60 + dep_time %% 100) %% 1440,
    sched_dep_time_continuous = (sched_dep_time %/% 100 * 60 +
      sched_dep_time %% 100) %% 1440,
    dep_delay_diff = dep_delay - dep_time_continuous + sched_dep_time_continuous
  )
select(flights_deptime, dep_delay_diff, everything())

filter(flights_deptime, dep_delay_diff != 0)
```
I expect the relation between these three variable should be dep_delay = dep_time - sched_dep_time. Then I compute the dep_delay with continuous number for dep_time and sched_dep_time. It turns out with that there are still many flights with none zero dep_delay_diff. All their dep_delay_diff equal to 1440 so we may know that all this flights were scheduled to depart before midnight but delay to the next day.

##  Exercise 4
```{r}
#
flights_delayed <- mutate(flights, 
                          dep_delay_min_rank = min_rank(desc(dep_delay)),
                          dep_delay_dense_rank = dense_rank(desc(dep_delay))
                          )
flights_delayed <- filter(flights_delayed, 
                          dep_delay_min_rank <= 10 ,dep_delay_dense_rank <=10)
select(arrange(flights_delayed, dep_delay_min_rank), dep_delay_min_rank, dep_delay_dense_rank, everything())
  

```