---
title: "Homework 5"
author: "Zeyan Huang"
date: "`r format(Sys.time(), '%B %d, %Y')`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Instructions

**Exercises:**  1,3 (Pg. 227); 2 (Pg. 232); 1,2 (Pg. 235); 

**Submission:** Submit via an electronic document on Sakai. Must be submitted as a HTML file generated in RStudio. All assigned problems are chosen according to the textbook *R for Data Science*. You do not need R code to answer every question. If you answer without using R code, delete the code chunk. If the question requires R code, make sure you display R code. If the question requires a figure, make sure you display a figure. A lot of the questions can be answered in written response, but require R code and/or figures for understanding and explaining.

```{r, include=FALSE}
library(tidyverse)
```

# Chapter 12 (Pg. 227)

##  Exercise 1
```{r}
#
rincome_plot <-
  gss_cat %>%
  ggplot(aes(x = rincome)) +
  geom_bar()
rincome_plot

rincome_plot +
  coord_flip()
```
The problem with default bar chart settings is that the labels overlap and are not easy to read. To improve this question, we can try to flip the axes. But we can also try to adjust the angle of the x-axis labels to some way that let them will not overlap and is easy to read.

##  Exercise 3
```{r}
#
gss_cat
levels(gss_cat$denom)

gss_cat %>%
  filter(!denom %in% c(
    "No answer", "Other", "Don't know", "Not applicable",
    "No denomination"
  )) %>%
  count(relig)

gss_cat %>%
  count(relig, denom) %>%
  ggplot(aes(x = relig, y = denom, size = n)) +
  geom_point() +
  theme(axis.text.x = element_text(angle = 90))
```

# Chapter 12 (Pg. 232)

##  Exercise 2
```{r}
#
keep(gss_cat, is.factor) %>% names()

levels(gss_cat$marital)

levels(gss_cat$race)
gss_cat %>%
  ggplot(aes(race)) +
  geom_bar()

levels(gss_cat$rincome)

levels(gss_cat$partyid)

levels(gss_cat$relig)
gss_cat %>%
  ggplot(aes(relig)) +
  geom_bar() +
  coord_flip()

levels(gss_cat$denom)
gss_cat %>%
  ggplot(aes(denom)) +
  geom_bar() +
  coord_flip()

```

The ordering of marital is “somewhat principled”. There is some sort of logic in that the levels are grouped “never married”, married at some point (separated, divorced, widowed), and “married”.

The ordering of race is principled in that the categories are ordered by count of observations.

The levels of rincome are ordered in decreasing order of the income but the placement of “No answer”, “Don’t know”, “Refused”, and “Not applicable” is arbitrary.

The levels of partyid are ordered from “Strong Republican”" to “Strong Democrat” besides the choices of “No answer”, “Don’t know”, and “Other party”.

The levels of relig and denom are arbitrary because there are no natural ordering, and they don’t appear to be ordered by stats within the dataset.
# Chapter 12 (Pg. 235)

##  Exercise 1
```{r}
#
levels(gss_cat$partyid)

gss_cat %>%
  mutate(
    partyid =
      fct_collapse(partyid,
        other = c("No answer", "Don't know", "Other party"),
        rep = c("Strong republican", "Not str republican"),
        ind = c("Ind,near rep", "Independent", "Ind,near dem"),
        dem = c("Not str democrat", "Strong democrat")
      )
  ) %>%
  count(year, partyid) %>%
  group_by(year) %>%
  mutate(proportion = n / sum(n)) %>%
  ggplot(aes(
    x = year, y = proportion,
    color = fct_reorder2(partyid, year, proportion)
  )) +
  geom_point() +
  geom_line() +
  labs(color = "Party")
```

##  Exercise 2
```{r}
#
levels(gss_cat$rincome)

gss_cat %>%
  mutate(
    rincome =
      fct_collapse(
        rincome,
        `Unknown` = c("No answer", "Don't know", "Refused", "Not applicable"),
        `Lt $5000` = c("Lt $1000", str_c(
          "$", c("1000", "3000", "4000"),
          " to ", c("2999", "3999", "4999")
        )),
        `$5000 to 10000` = str_c(
          "$", c("5000", "6000", "7000", "8000"),
          " to ", c("5999", "6999", "7999", "9999")
        )
      )
  ) %>%
  ggplot(aes(x = rincome)) +
  geom_bar() +
  coord_flip()
```
