---
title: "Homework 4"
author: "Zeyan Huang"
date: "`r format(Sys.time(), '%B %d, %Y')`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Instructions

**Exercises:**  2,3 (Pg. 151); 2,4 (Pg. 156); 1,2 (Pgs. 160-161); 2 (Pg. 163); 2,3,4 (Pg. 168)

**Submission:** Submit via an electronic document on Sakai. Must be submitted as a HTML file generated in RStudio. All assigned problems are chosen according to the textbook *R for Data Science*. You do not need R code to answer every question. If you answer without using R code, delete the code chunk. If the question requires R code, make sure you display R code. If the question requires a figure, make sure you display a figure. A lot of the questions can be answered in written response, but require R code and/or figures for understanding and explaining.

```{r, include=FALSE}
library(tidyverse)
```

# Chapter 9 (Pg. 151)

##  Exercise 2
```{r}
#
t2_cases <- filter(table2, type == "cases") %>%
  rename(cases = count) %>%
  arrange(country, year)
t2_population <- filter(table2, type == "population") %>%
  rename(population = count) %>%
  arrange(country, year)
t2_cases_per_cap <- tibble(
  year = t2_cases$year,
  country = t2_cases$country,
  cases = t2_cases$cases,
  population = t2_population$population
) %>%
  mutate(cases_per_cap = (cases / population) * 10000) %>%
  mutate(type = "cases_per_cap") %>%
  rename(count = cases_per_cap)
t2_cases_per_cap

t2_cases_per_cap <- t2_cases_per_cap %>%
  select(country, year, type, count)
table2b <-
  bind_rows(table2, t2_cases_per_cap) %>%
  arrange(country, year, type, count)

table2b


```


```{r}
table4c <-
  tibble(
    country = table4a$country,
    `1999` = table4a[["1999"]] / table4b[["1999"]] * 10000,
    `2000` = table4a[["2000"]] / table4b[["2000"]] * 10000
  )
table4c

```

If we only want to focus on the cases_per_cap, the second way which has the output table4c would be the easiest way to use. However, the output table2b by using the first way contains more information than table4c. Unfortunately, table2b is not tidy data. So, If we use the t2_cases_per_cap as the final output for the first way, it would be the most efficient way to see all data with separate columns for cases, population, and case_per_cap.

##  Exercise 3
```{r}
#
table2 %>%
  filter(type == "cases") %>%
  ggplot(aes(year, count)) +
  geom_line(aes(group = country), colour = "grey50") +
  geom_point(aes(colour = country)) +
  scale_x_continuous(breaks = unique(table2$year)) +
  ylab("cases")
```

The first thing we need to do is to filter the table to only include rows representing cases of TB.
# Chapter 9 (Pg. 156)

##  Exercise 2
```{r}
#
table4a %>%
 gather('1999', '2000', key = "year", value = "cases")
```
The code fail because 1999 and 2000 were not considered as names of columns but as integers here. If we want to note that we are mentioning the column names with numbers, we should use '' to emphasize.

##  Exercise 4
```{r}
#
preg <- tribble(
 ~pregnant, ~male, ~female,
 "yes", NA, 10,
 "no", 20, 12
)

table_preg <- preg %>%
  gather(sex, count, male:female) %>%
  drop_na()

table_preg
```

I choose to use gather(). The variables are about the status of pregnancy, the sex, and the count of people.

# Chapter 9 (Pgs. 160-161)

##  Exercise 1
```{r}
#
tibble(x = c("a,b,c", "d,e,f,g", "h,i,j")) %>%
 separate(x, c("one", "two", "three"))

tibble(x = c("a,b,c", "d,e,f,g", "h,i,j")) %>%
  separate(x, c("one", "two", "three"), extra = "merge")

tibble(x = c("a,b,c", "d,e,f,g", "h,i,j")) %>%
  separate(x, c("one", "two", "three"), extra = "drop")

tibble(x = c("a,b,c", "d,e", "f,g,i")) %>%
 separate(x, c("one", "two", "three"))

tibble(x = c("a,b,c", "d,e", "f,g,i")) %>%
  separate(x, c("one", "two", "three"), fill = "right")

tibble(x = c("a,b,c", "d,e", "f,g,i")) %>%
  separate(x, c("one", "two", "three"), fill = "left")
```

The extra argument tells separate() what to do if there are too many pieces of value, and the fill argument tells it what to do if there aren’t enough pieces of value. By default, separate() drops extra values with a warning. 
By setting extra = "merge", the extra piece would be combine into the last column. By setting extra = "drop", the R will output the same data as default but without the warning. 
By setting fill = "right", the function will fill with missing values from the right. By setting fill = "left", the function will fill with missing values from the left. With sepecific fill argument, the warning will not appear.

##  Exercise 2
```{r}
#
```

If the remove argument set as TRUE, it remove input column from output data frame.

# Chapter 9 (Pg. 163)

##  Exercise 2
```{r}
#
treatment <- tribble(
 ~ person, ~ treatment, ~response,
 "Derrick Whitmore", 1, 7,
 NA, 2, 10,
 NA, 3, 9,
 "Katherine Burke", 1, 4
)

treatment %>%
 fill(person, .direction = "up" )

```


With fill, the direction determines whether NA values should be replaced by the previous non-missing value ("down") or the next non-missing value ("up").

# Chapter 9 (Pg. 168)

##  Exercise 2
```{r}
#
who1 <- who %>%
 gather(
 new_sp_m014:newrel_f65, key = "key",
 value = "cases",
 na.rm = TRUE
 )

who3 <- who1 %>%
 separate(key, c("new", "type", "sexage"), sep = "_")

filter(who3, new == "newrel") %>% head()
```

The separate() function emits the warning “too few values”. If we check the rows for keys beginning with "newrel_", we see that sexage column is missing, and type = m014.

##  Exercise 3
```{r}
#
select(who3, country, iso2, iso3) %>%
  distinct() %>%
  group_by(country) %>%
  filter(n() > 1)


```

If the iso2 and iso3 are redundant with country, I expect that there are only one distinct combination of iso2 and iso3 value. And the output shows there are 0 country with more than one distinct combination of iso2 and iso3 value.

##  Exercise 4
```{r}
#
whoA <- who%>%
 gather(code, value = "cases", new_sp_m014:newrel_f65, na.rm = TRUE) %>%
 mutate(code = stringr::str_replace(code, "newrel", "new_rel")) %>%
 separate(code, c("new", "var", "sexage")) %>%
 select(-new, -iso2, -iso3) %>%
 separate(sexage, c("sex", "age"), sep = 1)

whoA %>%
  group_by(country, year, sex) %>%
  filter(year > 1995) %>%
  summarise(cases = sum(cases)) %>%
  unite(country_sex, country, sex, remove = FALSE) %>%
  ggplot(aes(x = year, y = cases, group = country_sex, colour = sex)) +
  geom_line()
```
